---
title: "Interactive Visualizations"
subtitle: "For Assignment 5"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
        code_folding: hide
---

```{r setup, message=FALSE, echo=FALSE, include=FALSE, eval=TRUE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)
library(mapboxapi)
library(leaflet)
library(sf)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 8,
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

```{css, echo = FALSE}
.code-r { /* Code block */
  font-size: 15px;
}

.code-r-small { /* Code block */
  font-size: 10px;
}
```



```{r load-data, include=FALSE, eval=TRUE}
source("resources/data/process_data.R")
```

Here are some interactive visualizations that showcase the occurrence of crime in different neighborhoods over time based on the observations.

<br>


## Mapview of Crime Occurrence {.tabset}

### Neighbourhood Scatters
<!-- 
https://plotly.com/r/scattermapbox/
https://plotly-r.com/maps.html
https://rstudio-pubs-static.s3.amazonaws.com/990225_5ed3a88ab43248dca24eaa8d4ecda370.html 
-->
```{r mapview1, class.source='foldable'}
# Group crime data by neighborhood
crime_by_neighb <- crime_data %>% 
  group_by(hood_id) %>% 
  summarise(
    longitude = mean(longitude),
    latitude = mean(latitude),
    neighbourhood = first(neighbourhood),
    counts = length(unique(event_unique_id))
  )

# Read in Mapbox access token
MAPBOX_TOKEN <- paste(readLines("resources/files/.mb_token"), collapse="") 
Sys.setenv('MAPBOX_TOKEN' = MAPBOX_TOKEN)

# Plot mapview
plot_mapbox(
    mode = 'scattermapbox'
  ) %>%
  add_markers(
    data = crime_by_neighb,
    y = ~ latitude,
    x = ~ longitude,
    size = ~ counts,
    sizemode = "area",
    color = ~ neighbourhood,
    hoverinfo = "text",
    text = ~ paste(neighbourhood, ":", counts, "times")
  ) %>% 
  layout(
    title = list(
      text = "Mapview of Crime Frequency in Neighborhoods",
      font = list(size = 16, color = "white"),
      y = 0.95, x = 0.5,
      xanchor = 'center', yanchor = 'top'
    ),
    mapbox = list(
      style = 'dark', zoom = 9,
      center = list(lat = median(crime_by_neighb$latitude),
                    lon = median(crime_by_neighb$longitude))
    ),
    # legend = list(orientation = 'h'),
    showlegend = FALSE
  ) 

```

### Neighbourhood Polygons
<!-- https://rstudio.github.io/leaflet/popups.html -->

```{r mapview2, class.source='foldable'}
# Generate geometry point for each obs
crime_filtered <- crime_data %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

# Aggregate data by neighborhood
crime_by_neighb_sf <- st_join(crime_filtered, neighbor_data, join = st_intersects) %>%
  group_by(AREA_NAME) %>%
  summarize(counts = length(unique(event_unique_id)))

# Merge neighborhood_data and crime_data
crime_by_neighb_sf <- st_join(neighbor_data, crime_by_neighb_sf)

# Create a leaflet map
pal <- colorNumeric(palette = "Blues", domain = crime_by_neighb_sf$counts)
crime_by_neighb_sf %>% 
  leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(fillColor = ~pal(counts),
              fillOpacity = .6,
              color = "grey",
              weight = .5,
              label = ~paste(AREA_NAME.x, ":", counts, "times"),
              labelOptions = labelOptions(
                style = list("font-size" = "12px")
              )) %>%
  addLegend(position = "bottomright", 
            title = "Number of Crimes",
            pal = pal, values = crime_by_neighb_sf$counts) 
```





## {- .unlisted}


- In previous observations, it was noted that shooting and firearm discharge incidents tend to be concentrated around the outskirts of the city, while auto theft incidents occur throughout Toronto but are more prevalent in the northwest region. On the other hand, break-and-enter and robbery incidents are predominantly concentrated in the downtown area of Toronto, where population density is relatively high. 

- These patterns can be observed from the map polygons, with the darkest shade indicating the highest number of crimes in the northwest region due to auto theft, and a relatively higher number of crimes in the downtown area of Toronto. In contrast, areas with low population density show lighter shade on the map indicating fewer crimes. 

- The scatter plots provide a more accurate representation of the average location of all crimes that occurred in each neighborhood area. The size of the scatter plots corresponds to the number of crimes, with larger scatter plots indicating higher crime counts. Similarly, larger scatter plots are observed in the northwest region and downtown area, indicating higher crime levels in these areas.




```{r include=FALSE}
library(mapboxapi)
mb_access_token(MAPBOX_TOKEN)
library(leaflet)

walk_5min <- mb_isochrone("City of Toronto",
                          profile = "driving",
                          time = 5,
                          
                          access_token = MAPBOX_TOKEN)

leaflet(walk_5min) %>%
  addMapboxTiles(style_id = "streets-v11",
                 username = "mapbox",
                 
                 access_token = MAPBOX_TOKEN) %>%
  addPolygons()


```

<br><hr><br><br>

## Crime Occurrence on Date 

```{r class.source = 'foldable'}
# Aggregated monthly crime occurrence by crime_cat
crime_by_mon_cat <- crime_data %>% 
  mutate(month = ym(paste0(year, '-', month_num))) %>% 
  group_by(month, crime_cat) %>% 
  summarise(
    counts = length(unique(event_unique_id))
  )

# Plot
moncatplot <- crime_by_mon_cat %>% 
  ggplot(aes(x = month, y = counts)) +
  geom_line(aes(color = crime_cat)) +
  theme_minimal() +
  labs(title = "Crime Occurrence for each Crime Category", 
       x = "Month", y = "Number of Crime", color = "Crime Category")

ggplotly(moncatplot) 
```



- Based on the plotted data for crime occurrences in different categories between 2014 and 2022, it is evident that auto theft has shown a clear increasing trend over time, while robbery has shown a decreasing trend. The number of shooting and firearm discharge incidents appears to remain relatively constant over time, although this category has fewer observations compared to the other three categories. Break and Enter, on the other hand, has the highest number of observations in the combined data, almost equal to the sum of the number of occurrences in the other three categories before the year 2018, but it shows a decreasing trend after that.


<br><hr><br><br>

## Crime Occurrence with Rates

<!-- https://plotly.com/r/multiple-axes/ -->

```{r class.source = 'foldable'}
# Combine rates_data with aggregated monthly crime occurrence
crime_by_mon <- crime_data %>% 
  mutate(month = ym(paste0(year, '-', month_num))) %>% 
  group_by(month) %>% 
  summarise(
    counts = length(unique(event_unique_id))
  ) %>% 
  inner_join(rate_data, by=c("month" = "month_date"))

# Plot the bar+lines 
crime_by_mon %>% 
  plot_ly(
    x = ~ month
  ) %>% 
  add_bars(
    y = ~ counts, name = "Number of Crimes", 
    marker = list(color = 'lightblue'),
    hoverinfo = "text",
    text = ~paste(
      month.abb[month.y], year, "\n", "Num of Crimes %:", counts
    )
  ) %>%
  add_lines(
    y = ~ inflation, name = "Inflation Rate", yaxis = "y2",
    hoverinfo = "text",
    text = ~paste(
      month.abb[month.y], year, "\n", "Inflation %:", inflation
    )
  ) %>% 
  add_lines(
    y = ~ unemployment, name = "Unemployment Rate", yaxis = "y2",
    hoverinfo = "text",
    text = ~paste(
      month.abb[month.y], year, "\n", "Unemployment %:", unemployment
    )
  ) %>% 
  layout(
    hovermode = "x",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Number of Crimes", side = "left"),
    yaxis2 = list(
      overlaying = "y",
      side = "right",
      title = "Rates (%)"
    ),
    title = list(
      text = "Crime Occurrence vs Inflation/Unemployment Rates",
      font = list(size = 18),
      y = 0.95, x = 0.5, xanchor = 'center', yanchor = 'top'
    ),
    legend = list(x = 1.1, y = .8)
  ) 
```

```{r include=FALSE}
# Compare inflation and unemployment rate
rate_data %>% 
  plot_ly(
    x = ~ month_date,
    hoverinfo = "text",
    text = ~ paste(
      month.abb[month], year, "\n\t",
      "Inflation %: ", inflation, "\n\t",
      "Unemployment %: ", unemployment
    )
  ) %>% 
  add_lines(y = ~ inflation, name = "Inflation") %>% 
  add_lines(y = ~ unemployment, name = "Unemployment") %>% 
  layout(
    hovermode = "compare",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Rate", side = "left"),
    title = list(
      text = "Inflation Rate vs Unemployment Rate",
      font = list(size = 20),
      y = 0.95, x = 0.5, xanchor = 'center', yanchor = 'top'
    ),
    legend = list(
      title=list(text='Rate'),
      y = 0.68
    )
  )


# Method 2: ggplotly()
count_by_month_cat <- crime_data %>% 
  group_by(year, month_num, crime_cat) %>% 
  summarise(
    count = length(unique(event_unique_id))
  )

barlineplt <- count_by_month_cat %>% 
  mutate(
    month = ym(paste0(year, '-', month_num))
  ) %>% 
  group_by(month) %>% 
  summarise(
    count = sum(count)
  ) %>% 
  inner_join(rate_data, by=c("month" = "month_date")) %>% 
  ggplot(aes(x = month, y = count)) +
  geom_bar(stat = "identity", fill = "lightblue")+
  theme_classic()+
  geom_line(aes(x = month, y = inflation * 100, group = 1, color = "Inflation")) +
  geom_point(aes(x = month, y = inflation * 100, color = "Inflation"), size = 2) +
  geom_line(aes(x = month, y = unemployment * 100, group = 1, color = "Unemployment")) +
  geom_point(aes(x = month, y = unemployment * 100, color = "Unemployment"), size = 2) +
  scale_color_manual(values = c("Inflation" = "blue", "Unemployment" = "red"), 
                     name = "Rates", labels = c("Inflation", "Unemployment")) +
  labs(
    title = "Number of Crime Occurrence over Time",
    x = "Year", y = "Number of Crimes"
  ) +
  scale_y_continuous(
    name = "Number of Crimes",
    sec.axis = sec_axis(~ . / 100, name = "Rates (%)" )
  )

ggplotly(barlineplt)
```


- The occurrence of crime in Toronto peaked at over 1500 occurrences per month in October 2019 between 2018 and 2022, but then decreased in the following two years to below 1300 occurrences per month. There appears to be a seasonal trend with lower crime rates in winter.

- Historically, inflation and unemployment have shown an inverse relationship, where inflation tends to decrease when unemployment rises. In other words, increased employment leads to higher spending power and demand. This pattern can be observed from the rate lines in the visualizations.

- According to the barplot depicting the number of crime occurrences, there appears to be a correlation between decreasing inflation rates around 2020 and a decrease in the number of crimes. Additionally, starting from 2021, as the inflation rate continues to increase, the number of crimes also decreases.

- Despite the sharp increase in the unemployment rate at the beginning of 2020 due to COVID-19, it appears to have had no discernible effect on the number of crimes committed. Moreover, in later 2021, despite a decrease in the unemployment rate, the number of crime occurrences continued to increase. This suggests that the unemployment rate may not be a significant factor in the occurrence of crimes.


<br><br><br>